---
title: "Improved Code: Exercise_Set 1"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org/"))

# Load necessary packages
#install.packages("tidyverse")
library(tidyverse)

#install.packages("ggplot2")
library(ggplot2)

 #install.packages("kableExtra")
library(kableExtra)

```

# Problem 1: Exploring Anscombe’s Quartet

## Understanding the Dataset
Anscombe’s Quartet demonstrates how summary statistics can be misleading and highlights the importance of data visualization. 

```{r load_anscombe}
# Load dataset
anscombe_quartet <- readRDS("data/anscombe quartet.rds")

# Inspect the dataset structure
str(anscombe_quartet)
```

**What does `str()` do?**
The `str()` function displays the internal structure of an R object, providing an overview of its type, dimensions, and data format.

## Summary Statistics
```{r summary_anscombe}
summary_table <- anscombe_quartet %>% 
  group_by(dataset) %>% 
  summarise(
    mean_x = mean(x),
    mean_y = mean(y),
    min_x = min(x),
    min_y = min(y),
    max_x = max(x),
    max_y = max(y),
    correlation = cor(x, y)
  )

# Display summary statistics as a nicely formatted table
summary_table %>% kable(caption = "Summary Statistics of Anscombe’s Quartet") %>% kable_styling()
```

**What do the summary statistics tell us?**
Despite having nearly identical means, variances, and correlations, the datasets are structurally very different. This emphasizes the importance of visualization.

## Visualizing the Data
```{r plot_anscombe}
anscombe_plot <- ggplot(anscombe_quartet, aes(x = x, y = y)) +
  geom_point(color = "blue", size = 2) + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "red") +
  facet_wrap(~dataset) +
  theme_minimal() +
  labs(title = "Anscombe’s Quartet Visualization", x = "X Values", y = "Y Values")

print(anscombe_plot)
```

## Saving the Plot
```{r save_anscombe, eval=TRUE}
ggsave("anscombe_quartet.png", anscombe_plot, width = 5, height = 5, dpi = 300)
```

**Interpretation of Plots:**
- Each dataset has the same summary statistics but vastly different distributions.
- Some show non-linear trends, while others have outliers that influence correlations.
- This reinforces why **visualizing data** is crucial before making conclusions.

**Would linear regression be appropriate?**
- In some cases, yes (e.g., dataset 1), but others display non-linearity or outliers that violate regression assumptions.

# Problem 2: Exploring the Datasaurus Dozen

## Load and Inspect Data
```{r load_datasaurus}
datasaurus_dozen <- readRDS("datasaurus_dozen.rds")

# Structure of the dataset
str(datasaurus_dozen)
```

## Summary Statistics for Datasaurus Dozen
```{r summary_datasaurus}
datasaurus_summary <- datasaurus_dozen %>% 
  group_by(dataset) %>% 
  summarise(
    mean_x = mean(x),
    mean_y = mean(y),
    min_x = min(x),
    min_y = min(y),
    max_x = max(x),
    max_y = max(y),
    correlation = cor(x, y)
  )

# Display summary in a styled table
kable(datasaurus_summary, caption = "Summary Statistics of Datasaurus Dozen") %>% kable_styling()
```

## Visualizing the Different Datasets
```{r plot_datasaurus}
datasaurus_plot <- ggplot(datasaurus_dozen, aes(x = x, y = y)) +
  geom_point(color = "purple", alpha = 0.6) + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "orange") +
  facet_wrap(~dataset, scales = "free") +
  theme_minimal() +
  labs(title = "Datasaurus Dozen Visualization", x = "X Values", y = "Y Values")

print(datasaurus_plot)
```

## Save the Plot
```{r save_datasaurus, eval=TRUE}
ggsave("datasaurus_dozen.png", datasaurus_plot, width = 6, height = 6, dpi = 300)
```

**Observations:**
- Even though datasets have similar summary statistics, their visual patterns vary significantly.
- Some datasets form shapes (e.g., a dinosaur!), while others show distinct trends.
- Again, this highlights why **visualization matters** beyond summary statistics.

# Conclusion
- Summary statistics alone can be misleading; always visualize your data.
- Linear regression isn't always the best choice—check for patterns, outliers, and non-linear relationships.
- Graphs provide insights that numbers alone cannot!

---
This document is rendered as **HTML** and **PDF** to showcase results in different formats. 



